The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
More corrections to chapter 1 and 2.
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1737 4525493e-7705-40b1-a816-d608a930855b

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database(s)
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of only documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from
single table or joined tables. While \cite{QueryAnswer} make a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. Those and other works\footnote{
For more information about
keyword based search over relational databases we redirect users to reference in
\cite{DBXplorer, QueryAnswer}.} 
are based on some kind of the graph over the underlying database and
moreover use a generated database out of the pristine one.
The proximity of the results in those techniques 
are mostly confined to the text-based values stored in a database, since the
keyword based search with numerical values 
need the precise knowledge of the related entity in a schema. For example, typing
100 can lead to inappropriate matches, such as row ids, and in order to be
answered correctly it requires additional keyword, clarifying its context, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore such approaches cannot be used to answer the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100 taken from january to february of 2009} which can be easily accomplished
using SQL syntax.

To address those limitations an attempt to build a simple, intuitive and 
flexible query language was introduced in \cite{DBS-QL}\footnote{A similar but less
sophisticated approach was used in \cite{AMI}}.
It represents a power of SQL while hiding underlying relational schema from 
the end-users. As a results a human questions were intuitively mapped into 
simple queries. For example, the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
This solution represents a common use cases within HEP community and provide
intuitive mapping between mental model of end-users and underlying database back-end.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into federated DB where SQL queries can be placed to 
search desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated database it still requires 
an access to the underlying database schema. This obviously creates a problem
working in distributed heterogeneous environment where different security policies
are in place.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are defined by developers.
Keeping in mind that meta-data can come in different forms and data formats
we used a document-oriented database back-end with full support of query language.
The key ingredients were the setup of the mapping and analytics services which provide
necessary glue across existing API/notations and user queries. This approach 
allows us to perform the data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two large general-purpose particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of High Energy Physics and provide physicists
ability to look at conditions presented at early stage of our Universe.
More then 3000 physicists from 183 institutions reprepresenting almost 
38 countries are involved in the design, construction and maintenance of the experiment.

@@

\bibitem{AMI}
Altas AMI web portal,
http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html

>>>>

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database(s)
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of only documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. These and other related 
works\footnote{For more information about keyword based search over relational 
databases we redirect users to references in \cite{DBXplorer, QueryAnswer}.} 
are based on some kind of the graph over the underlying database and
moreover use a generated database out of the pristine one.
The proximity of the results in those techniques 
are mostly confined to the text-based values stored in a database, since 
usage of numerical values in the input search field is mostly useless without 
knowledge of the context of the value. For example, typing a value 100 
in the search input field leads to plenty of inappropriate matches, 
such as row ids, and in order to be
answered correctly requires additional keyword, clarifying its meaning, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore to answer presicely the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100} would be hard or almost impossible using keyword search techniques discussed 
in \cite{DBXplorer, QueryAnswer}, while it can be easily accomplished
using SQL syntax.

To address those limitations an attempt to build a simple, intuitive and 
flexible query language was introduced in \cite{DBS-QL}.
It represents a power of SQL while hiding underlying relational schema from 
the end-users. As a results a human questions were intuitively mapped into 
simple queries. For example, the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
This solution represents a common use cases within HEP community and provide
intuitive mapping between mental model of end-users and underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into the federated DB where SQL queries can be placed to 
search for desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated database it still requires 
an access to the underlying database schema. This obviously creates a problem
in distributed heterogeneous environment where different security policies
are in place.

To avoid those limitations we decided to build a keyword-search based system around
existing data-services whose nature and access policies are defined by developers.
Keeping in mind that meta-data can come in different forms and data formats
we used a document-oriented database back-end with full support of the query language.
The key ingredients were the setup of the mapping and analytics services which provide
necessary glue across existing API/notations and user queries. This approach 
allows us to perform the dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.
%to store meta-data information 
%We wanted to expand this approach further and apply it 
%to handle cross data-service queries from broad variety 
%of relational and non-relational data-sources within 
%our distributed heterogeneous environment.

\section{CMS data model\label{DataModel}}
The CMS, Compact Muon Solenoid, \cite{CMS} 
is one of the two large general-purpose particle physics detectors built on 
the proton-proton Large Hadron Collider (LHC) at CERN in Switzerland and France. 
It is designed to explore the frontier of the High Energy Physics and provide physicists
ability to look at conditions presented at the early stage of our Universe.
More then 3000 physicists from 183 institutions reprepresenting almost 
38 countries are involved in the design, construction and maintenance of the experiment.

@@

%\bibitem{AMI}
%Altas AMI web portal,
%http://ami.in2p3.fr/opencms/opencms/AMI/www/Tutorial/AMIMediation.html
